---
title: "Length of Stay Cox Regression Analysis"
output: html_notebook
---

```{r}
library(survival)
library(tidyverse)
library(ggfortify)
```

```{r}
#surv_data <- readRDS(file = "../data/ed_ex.rds")
#surv_data <- readRDS(file = "../../cw-data/cw_ip_spells.rds")
#surv_data <- readRDS(file = "../../lgt-data/data-out/spell_table_lgt_20190816_b704e260_ms.rds")
surv_data <- readRDS(file = "../../cw-data/spell_table_cw_20190902_b7fae6f7.rds")

surv_data <- hospitalflow::add_directorate_variable(surv_data)

surv_data <- surv_data %>% filter(admission_method_type == "Emergency Admissions",
                                  !(age_band_start %in% c("1-4 yrs", "5-9 yrs", "10-14 yrs")),
                                  directorate == "Medical"#,
                                  #main_specialty_start %in% c("ACUTE INTERNAL MEDICINE", "GENERAL MEDICINE")
                                  ) %>%
  mutate(start_datetime = if_else(is.na(initial_ed_end_datetime), spell_start, initial_ed_end_datetime),
         end_datetime = spell_end,
         age_band_start = forcats::fct_drop(age_band_start))
surv_data <- surv_data %>% mutate(los = difftime(end_datetime, start_datetime), adm_wd = weekdays(start_datetime)) %>% filter(!is.na(los)) %>%
  mutate(days_since_prev_disch = as.numeric(replace(days_since_prev_disch, days_since_prev_disch <= 0, NA)),
         disch_in_last_week = if_else(!is.na(days_since_prev_disch) & days_since_prev_disch <=7, TRUE, FALSE),
         disch_in_last_month = if_else(!is.na(days_since_prev_disch) & days_since_prev_disch <=30, TRUE, FALSE),
         admitted_before = !is.na(days_since_prev_disch))
units(surv_data$los) <- "hours"
clean_surv_data <- surv_data %>% filter(!is.na(los) & los >= 0 & end_datetime < lubridate::ymd_hm('2019-03-01 00:00'))


obsLen <- 365*24 # length of observation time (censoring time = end of study)
clean_surv_data <- clean_surv_data %>% mutate(eventT = los,
                                              obsT = pmin(eventT, obsLen),
                                              status = eventT <= obsLen,
                                              gender = dplyr::if_else(gender == "Male", 1, 0),
                                              adm_calmonth = as.factor(format(start_datetime, "%b")),
                                              adm_daytime = dplyr::if_else(dplyr::between(lubridate::hour(start_datetime), 8, 16), 1, 0))
# eventT <- clean_surv_data %>% pull(los)
# adm_wd <- clean_surv_data %>% pull(adm_wd)
# obsT   <- pmin(eventT, obsLen)  # observed censored event times
# status <- eventT <= obsLen     # has event occured?

dfSurv <- clean_surv_data %>% select(eventT, obsT, status)
dfSurv2 <- clean_surv_data %>% select(spell_number, eventT, obsT, status,
                                      adm_wd, adm_calmonth, gender, age_band_start,
                                      start_datetime, end_datetime, adm_daytime, days_since_prev_disch, disch_in_last_week, disch_in_last_month, admitted_before)
```

```{r}
ggplot(dfSurv, aes(eventT)) + stat_ecdf(geom = "step") +
  scale_x_continuous(breaks = scales::pretty_breaks(n=12), limits = c(0, 30L*24L)) +
  ggtitle("Cumulative time since arrival distribution") +
  xlab("Time from arrival (t)") + ylab("F(t)") +
  geom_vline(xintercept = obsLen, colour = "blue") #+
  #geom_vline(xintercept = 4, colour = "green", linetype = "longdash")
```

```{r}
KM0 <- survfit(Surv(obsT, status) ~ 1,  type="kaplan-meier", conf.type="log", data=dfSurv)
KM1 <- survfit(Surv(obsT, status) ~ adm_wd,  type="kaplan-meier", conf.type="log", data=dfSurv2)
```

```{r}
autoplot(KM1) + ggtitle(expression(paste("Kaplan-Meier estimate ", hat(S)(t), " with CI"))) +
  xlab("Time from arrival (t)") + ylab(expression(paste(hat(S)(t), ": % remaining in hospital"))) +
  geom_vline(xintercept = obsLen, colour = "blue") +
  scale_x_continuous(breaks = scales::pretty_breaks(n=12), limits = c(0, 30L*24L))
```

```{r}
S_function <- hospitalflow::make_survival_function(KM0)

cond_S_function <- function(t0, t1, S_function) {
  pmin(S_function(t1)/S_function(t0),1)
}

ggplot(data.frame(x=c(0,30L*24L)), aes(x)) +
  stat_function(fun = S_function, geom = "line", aes(colour = "S(x)")) +
  stat_function(fun = function(x) cond_S_function(1L*24L, x, S_function),
                geom = "line", aes(colour = "S(x|t0 = 1 day)")) +
  stat_function(fun = function(x) cond_S_function(2L*24L, x, S_function),
                geom = "line", aes(colour = "S(x|t0 = 2 days)")) +
  stat_function(fun = function(x) cond_S_function(4L*24L, x, S_function),
                geom = "line", aes(colour = "S(x|t0 = 4 days)")) +
  stat_function(fun = function(x) cond_S_function(7L*24L, x, S_function),
                geom = "line", aes(colour = "S(x|t0 = 7 days)")) +
  stat_function(fun = function(x) cond_S_function(14L*24L, x, S_function),
                geom = "line", aes(colour = "S(x|t0 = 14 days)")) +
  ggtitle("Probability of remaining in hospital after time t1, given in hospital after time t0") +
  xlab("Time from arrival (t) hours") + ylab("Probability still in hospital at time t") +
  scale_x_continuous(breaks = seq(0, 30L*24L, by = 24L), limits = c(0, 30L*24L)) +
  geom_vline(xintercept = obsLen, colour = "blue") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```


```{r}
T1 <- strptime("2018-09-01 12:00:00", format = "%Y-%m-%d %H:%M:%S", tz = "Europe/London")
```

# Predicting Residual Occupancy

```{r}

hospitalflow::predict_residual_occupancy(df = surv_data, t = T1, S_function = S_function, delta_t = 48)
hospitalflow::get_residual_occupancy(df = surv_data, t = T1, delta_t = 48)

dates <- seq(lubridate::ymd_hm('2018-08-01 12:00', tz = "Europe/London"), lubridate::ymd_hm('2019-01-31 12:00', tz = "Europe/London"), by = '1 day')

model_output <- hospitalflow::run_S_func_model(df = surv_data, S_func = S_function, date_seq = dates)

summary(model_output$predictions)
```

```{r}
model_output$diagnostics$comp_plot
```

```{r}
model_output$diagnostics$err_plot
```

```{r}
model_output$diagnostics$rel_err_plot
```


The errors clearly show (as expected) the effect of weekends on this model - this naive model does not take into account any weekend effect. 

Next steps will be to generalise the model to include this weekly effect.

## Cox PH Model

```{r coxph-model1}
CM1 <- coxph(Surv(obsT, status) ~ adm_wd, data = dfSurv2)
KM1S <- survfit(CM1, newdata = data.frame(adm_wd = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")))

step_funs <- lapply(c(1:7), function(x) {hospitalflow::make_survival_function(KM1S[,x])})
names(step_funs) <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")
```

```{r coxph-predict1}
ipss <- hospitalflow::get_inpatient_snapshot(df = dfSurv2, t = T1) %>% filter(!is.na(stay_duration))
newdata1 <- ipss %>% select(obsT = stay_duration, adm_wd) %>% mutate(status = TRUE)
newdata1d <- newdata1 %>% mutate(obsT = obsT + 48)
CM1_prediction_1 <- predict(CM1, newdata = newdata1, type = "expected")
CM1_prediction_2 <- predict(CM1, newdata = newdata1d, type = "expected")
prob_here_48h_CM1 <- exp(-CM1_prediction_2)/exp(-CM1_prediction_1)



predictions_CM1 <- cbind(ipss, CM1_pred_prob = prob_here_48h_CM1)

sum(predictions_CM1$CM1_pred_prob)
# Now lets embed that into predict_residual_occupancy

hospitalflow::predict_residual_occupancy(df = dfSurv2, t = T1, delta_t = 48,
                                         coxmodel = "Surv(obsT, status) ~ adm_wd")
hospitalflow::get_residual_occupancy(df = dfSurv2, t = T1, delta_t = 48)
```

```{r eval-coxph1}
model_output_CM <- hospitalflow::run_S_func_model(df = dfSurv2, coxmodel = "Surv(obsT, status) ~ adm_wd",
                                                  date_seq = dates)

summary(model_output_CM$predictions)
```

```{r}
model_output_CM$diagnostics$comp_plot
```

```{r}
model_output_CM$diagnostics$err_plot
```

```{r}
model_output_CM$diagnostics$rel_err_plot
```

```{r eval-coxph2}
CM2 <- coxph(Surv(obsT, status) ~ adm_wd + adm_daytime + adm_calmonth + gender + age_band_start, data = dfSurv2)
newdata2 <- ipss %>% select(obsT = stay_duration, adm_wd, adm_daytime, adm_calmonth, gender, age_band_start) %>% mutate(status = TRUE)
newdata2d <- newdata2 %>% mutate(obsT = obsT + 48)
CM2_prediction_1 <- predict(CM2, newdata = newdata2, type = "expected")
CM2_prediction_2 <- predict(CM2, newdata = newdata2d, type = "expected")
prob_here_48h_CM2 <- exp(-CM2_prediction_2)/exp(-CM2_prediction_1)

model_output_CM2 <- hospitalflow::run_S_func_model(df = dfSurv2,
                                                   coxmodel = "Surv(obsT, status) ~ adm_wd + adm_daytime + adm_calmonth + gender + age_band_start",
                                                  date_seq = dates)

summary(model_output_CM2$predictions)
```

```{r}
model_output_CM2$diagnostics$comp_plot
```

```{r}
model_output_CM2$diagnostics$err_plot
```

```{r}
model_output_CM2$diagnostics$rel_err_plot
```

## Cox model with previous admission variables added as covariates

```{r eval-coxph3}
CM3 <- coxph(Surv(obsT, status) ~ adm_wd + adm_daytime + adm_calmonth + gender + age_band_start + disch_in_last_week + disch_in_last_month + admitted_before,
             data = dfSurv2)
newdata3 <- ipss %>% select(obsT = stay_duration, adm_wd, adm_daytime, adm_calmonth, gender, age_band_start, disch_in_last_week, disch_in_last_month, admitted_before) %>% mutate(status = TRUE)
newdata3d <- newdata3 %>% mutate(obsT = obsT + 48)
CM3_prediction_1 <- predict(CM3, newdata = newdata3, type = "expected")
CM3_prediction_2 <- predict(CM3, newdata = newdata3d, type = "expected")
prob_here_48h_CM3 <- exp(-CM3_prediction_2)/exp(-CM3_prediction_1)

model_output_CM3 <- hospitalflow::run_S_func_model(df = dfSurv2,
                                                   coxmodel = "Surv(obsT, status) ~ adm_wd + adm_daytime + adm_calmonth + gender + age_band_start + disch_in_last_week + disch_in_last_month + admitted_before",
                                                  date_seq = dates)

summary(model_output_CM3$predictions)
```

```{r}
model_output_CM3$diagnostics$comp_plot
```

```{r}
model_output_CM3$diagnostics$err_plot
```

```{r}
model_output_CM3$diagnostics$rel_err_plot
```

## Naive model for comparison

```{r naive-ts}
naive_preds <- model_output_CM2$predictions %>% dplyr::select(dates, actual)

naive_preds <- naive_preds %>% dplyr::mutate(predicted = dplyr::lag(actual, 2)) %>% dplyr::filter(!is.na(predicted))

naive_preds <- naive_preds %>% 
  dplyr::mutate(error = predicted - actual, rel_error = error/actual,
                abs_error = abs(error), abs_rel_error = abs(rel_error))

prediction_comparison <- naive_preds %>% select(dates, predicted, actual) %>% gather(key = "type", value = "residual_occ", predicted, actual)
comp_plot <- ggplot2::ggplot(data = prediction_comparison, mapping = aes(dates, residual_occ)) + ggplot2::geom_point(aes(group = type, colour = type)) + geom_line(aes(group = type, colour = type)) + ggtitle("Residual occupancy: predictions versus actual") +
  xlab("Date of prediction") + ylab("Residual occupancy")

err_plot <- ggplot2::ggplot(data = naive_preds, mapping = aes(dates, error)) + ggplot2::geom_point() + geom_line() + ggtitle("Residual occupancy: errors in predictions versus actual") +
  xlab("Date of prediction") + ylab("Error")

rel_err_plot <- ggplot2::ggplot(data = naive_preds, mapping = aes(dates, rel_error)) + ggplot2::geom_point() + geom_line() + ggtitle("Residual occupancy: relative errors in predictions versus actual") +
  xlab("Date of prediction") + ylab("Relative Error")

summary(naive_preds)
```


```{r}
comp_plot
```

```{r}
err_plot
```

```{r}
rel_err_plot
```

AAU - more consistency 7 day
Downstream - much less so - could assume no weekend discharges?

General framework above does work to deploy and evaluate these models. However would be better to use the structure of e.g. caret!

This includes:
1. A structure to better represent (conditional) survival functions
2. A function or function template for calculating these survival functions from (a specified, and eventually possibly rolling, "training" subset of) the data
3. A function for applying these survival functions to the data at a given point in time to yield a delta-t horizon prediction for residual occupancy
4. A function or functions to evaluate these predictions across a "test" range

Idea: can I use the newdata argument of survfit to provide the actual covariates of snapshot data, to get a) S(t1 | covs) and b) S(t0 | covs), then compute their ratio to get the probability of still being in? If so this saves a lot of machinery in e.g. predict_residual_occupancy. - This is essentially what the Cox model approach above is doing.

Evaluation by occupancy "zones" - ROC?
Time series - esp. for delta on the predictions
v v v Busy ED/AAU : ?more admissions
Visualise to see the prediction before the effect
Time since prev adm- 
Existence of certain diagnoses previuos: chest pain, asthma, overdose, falls, found on floor, hf copd, dka
Prb of overfull given predicted residual occupancy + distrib of expected admits over 48hrs...

?incorporate lou's data?



**********
28th Jan 2020

- need to describe variance in res occ
- why are there so many emergency medical inpatients - 400 beds total? ~175/80 beds medicine? for monday snapshots to understand why 300 odd??
- plot survival curves by age-band 

- diagnostic prev
- new pts : poisson
- instances of change of specialty?
