---
title: "Length of Stay Survival Analysis"
output: html_notebook
---

Based on [this](http://dwoll.de/rexrepos/posts/survivalKM.html) article.

```{r}
library(survival)
library(tidyverse)
library(ggfortify)
```

```{r}
#surv_data <- readRDS(file = "../data/ed_ex.rds")
surv_data <- readRDS(file = "../../cw-data/cw_ip_spells.rds")
surv_data <- surv_data %>% mutate(los = difftime(end_datetime, start_datetime)) %>% filter(!is.na(los))
units(surv_data$los) <- "hours"
eventT <- surv_data %>% filter(!is.na(los) & los >= 0) %>% pull(los) 
obsLen <- 365*24                  # length of observation time (censoring time = end of study)
obsT   <- pmin(eventT, obsLen)  # observed censored event times
status <- eventT <= obsLen     # has event occured?
dfSurv <- data.frame(eventT, obsT, status)          # data frame
```

```{r}
ggplot(dfSurv, aes(eventT)) + stat_ecdf(geom = "step") +
  scale_x_continuous(breaks = scales::pretty_breaks(n=12), limits = c(0, 30L*24L)) +
  ggtitle("Cumulative time since arrival distribution") +
  xlab("Time from arrival (t)") + ylab("F(t)") +
  geom_vline(xintercept = obsLen, colour = "blue") #+
  #geom_vline(xintercept = 4, colour = "green", linetype = "longdash")
```

```{r}
KM0 <- survfit(Surv(obsT, status) ~ 1,  type="kaplan-meier", conf.type="log", data=dfSurv)
```

```{r}

autoplot(KM0) + ggtitle(expression(paste("Kaplan-Meier estimate ", hat(S)(t), " with CI"))) +
  xlab("Time from arrival (t)") + ylab(expression(paste(hat(S)(t), ": % remaining in hospital"))) +
  scale_x_continuous(breaks = scales::pretty_breaks(n=12), limits = c(0, obsLen*1.1)) +
  geom_vline(xintercept = obsLen, colour = "blue") +
  scale_x_continuous(breaks = scales::pretty_breaks(n=12), limits = c(0, 30L*24L)) #+
  #geom_vline(xintercept = 4, colour = "green", linetype = "longdash")
```

```{r}
KM0t <- tibble(time = KM0$time, n.risk = KM0$n.risk, n.event = KM0$n.event, n.censor = KM0$n.censor, surv = KM0$surv)
```

```{r}
cond_surv_from_km <- function(t0, KM) {
t1 <- c(t0:(30L*24L))
St0 <- KM %>% filter(time == t0) %>% pull(surv)
condSurv <- tibble(t1 = t1)
condSurv <- condSurv %>% left_join(KM %>% select(time, surv), by = c("t1" = "time")) %>%
  mutate(condSurvival = surv/St0)
condSurv %>% mutate(t0 = t0)
}

condSurvs <- lapply(seq(24L, 24L*7L, by = 24L), cond_surv_from_km, KM0t)

condSurv <- dplyr::bind_rows(condSurvs)

ggplot(condSurv, aes(t1, condSurvival)) + geom_line(aes(group = factor(t0))) + ggtitle("Probability of remaining in hospital after time t1, given in hospital after time t0 = ") +
  xlab("Time from arrival (t)") + ylab("Probability still in hospital at t1") +
  scale_x_continuous(breaks = seq(0, 30L*24L, by = 24L), limits = c(0, 30L*24L)) +
  geom_vline(xintercept = obsLen, colour = "blue") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))#+
  #geom_vline(xintercept = 4, colour = "green", linetype = "longdash")
```

```{r}
#t0 <- c(60,120,180,240,300)
#t1 <- c(61:(60*24))
#St0 <- KM0t %>% filter(time == t0) %>% pull(surv)
```

Next we should estimate the variability of this for typical medical population e.g. 180 medical pts.
H: with full info can predict e.g. 3day horizon given 2day; but not for e.g. 100days.

So lets say we have 30 (or 300) patients who have been in hospital already for (random exponentially distributed) times (in hours):
```{r}
times_in_hospital_1 <- floor(rexp(30, 1/51.6))
times_in_hospital_2 <- floor(rexp(300, 1/51.6))
times_in_hospital_1
times_in_hospital_2
```
Now use the survival function from above to estimate the probability each patient will still be in after a further 48 hours.
```{r}
probabilities_still_in_48h_1 <- sapply(times_in_hospital_1, hospitalflow::get_discharge_probability, KM0t, 48)
probabilities_still_in_48h_1
probabilities_still_in_48h_2 <- sapply(times_in_hospital_2, hospitalflow::get_discharge_probability, KM0t, 48)
probabilities_still_in_48h_2
```
These are the parameters of individual bernoulli trials, the sum of which yields a Poisson Binomial distribution. It is possible to calculate this distribution via the discrete Fourier transform of the following vector:
```{r}
p1 <- probabilities_still_in_48h_1
x1 <- (1/(length(p1)+1)) * hospitalflow::get_poisson_binomial_characteristic_values(p1)
p2 <- probabilities_still_in_48h_2
x2 <- (1/(length(p2)+1)) * hospitalflow::get_poisson_binomial_characteristic_values(p2)
```
The Poisson Binomial distribution mass function is now given by (for the sample of 30):
```{r}
pb_dist1 <- fft(x1)
pb_dist1 <- tibble::tibble(prob = Re(pb_dist1)) %>% tibble::rowid_to_column("k") %>% mutate(k = k-1)
expected_k1 <- pb_dist1 %>% dplyr::mutate(prod = prob*k) %>%
  summarise(expectation = sum(prod)) %>% pull(expectation)
ggplot(data = pb_dist1, aes(x = k, y = prob)) + geom_bar(stat = "identity") + geom_vline(xintercept = expected_k1, colour = "green")
expected_k1
```

The Poisson Binomial distribution mass function is now given by (for the sample of 300):
```{r}
pb_dist2 <- fft(x2)
pb_dist2 <- tibble::tibble(prob = Re(pb_dist2)) %>% tibble::rowid_to_column("k") %>% mutate(k = k-1)
expected_k2 <- pb_dist2 %>% dplyr::mutate(prod = prob*k) %>%
  summarise(expectation = sum(prod)) %>% pull(expectation)
ggplot(data = pb_dist2, aes(x = k, y = prob)) + geom_bar(stat = "identity") + geom_vline(xintercept = expected_k2, colour = "green")
expected_k2
```


Next step is to write a function to run this on patients actually in hospital/ward at a specified time t.

```{r}

```

