---
title: "Example Hospital Flow Report"
author: "CLAHRC NWL Information Intelligence Team"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(magrittr)

sample_duration <- lubridate::duration(4, "weeks")
```

Package version: `r stringr::str_sub(system("git rev-parse HEAD", intern=TRUE), 1, 8)`

```{r import-data, cache=TRUE, echo=FALSE}
# Import and Standardise Data

import_list <- readRDS("../lgt-config/lgt_import_list.rds")

# This import list uses relative paths (good for us during development)
# which means the paths need prepending with "../" to work from the
# working directory used in running this file.
# In production use, users will be best specifying their import paths
# as absolute paths.
import_list <- lapply(import_list, function(x) {
  list(data_path = paste0("../", x$data_path),
  config_path = paste0("../", x$config_path))
})

data_list <- hospitalflow::import_and_standardise(import_list)
```

```{r spell-table, cache=TRUE, echo=FALSE}
ed_data <- data_list[[1]]
inpatient_data <- data_list[[3]]

data_start_date <- as.Date(max(min(ed_data$start_datetime, na.rm = TRUE),
                       min(ed_data$end_datetime, na.rm = TRUE),
                       min(inpatient_data$start_datetime, na.rm = TRUE),
                       min(inpatient_data$end_datetime, na.rm = TRUE)))
data_end_date <- as.Date(min(max(ed_data$start_datetime, na.rm = TRUE),
                     max(ed_data$end_datetime, na.rm = TRUE),
                     max(inpatient_data$start_datetime, na.rm = TRUE),
                     max(inpatient_data$end_datetime, na.rm = TRUE)))
start_date <- data_start_date
end_date <- data_end_date

if(!is.na(sample_duration)) {
  
  data_start_date <- lubridate::ceiling_date(data_start_date, "month")
  data_end_date <- lubridate::floor_date(data_end_date, "month")
  
  if(data_end_date - data_start_date > sample_duration) {
    start_date <- data_end_date - sample_duration
    end_date <- data_end_date
    example_ed_data <- ed_data %>% dplyr::filter(start_datetime <= end_date,
                                                 end_datetime >= start_date)
    example_inpatient_data <- inpatient_data %>% dplyr::filter(start_datetime <= end_date,
                                                               end_datetime >= start_date)
  } else {
    example_ed_data <- ed_data
    example_inpatient_data <- inpatient_data
  }
} else {
  example_ed_data <- ed_data
  example_inpatient_data <- inpatient_data
}

example_spell_table <- hospitalflow::make_spell_table(ed_data = example_ed_data,
                                                      inpatient_data = example_inpatient_data)
```
## Example dataset
This example report is based on `r nrow(example_ed_data)` emergency department attendances and `r nrow(example_inpatient_data)` inpatient episodes of care, linked into `r nrow(example_spell_table)` hospital spells. All attendances and admissions occurred prior to `r end_date` and all discharges after `r start_date`.

## Age and Gender

```{r age-gender-ed, echo=F, results='asis', fig.height=5, fig.width=7, cache=TRUE}
hospitalflow::ae_attendances_admissions_age_sex(start_date = start_date, end_date = end_date,
                                                data = example_spell_table,
                                                plot_chart = TRUE,
                                                hospital_name = "Anytown General Hospital")
```

## Arrival Occupancy

```{r arr-occ, echo=F, results='asis', fig.height=5, fig.width=7, cache=TRUE}
hospitalflow::ae_arrival_occupancy(start_date = as.POSIXct(start_date, tz = "Europe/London"),
                                   end_date = as.POSIXct(end_date, tz = "Europe/London"),
                                   data = example_spell_table,
                                   plot_chart = TRUE,
                                   hospital_name = "Anytown General Hospital")
```

## Duration in ED by admitted/non-admitted

```{r los-ed, echo=F, results='asis', fig.height=5, fig.width=7, cache=TRUE}
hospitalflow::los_att_adm_ae(start_date = start_date, end_date = end_date,
                             data = example_spell_table,
                             plot_chart = TRUE,
                             hospital_name = "Anytown General Hospital")
```


## Admissions and Discharges by Day of Week


```{r adm-disch-dow, echo=F, results='asis', fig.height=5, fig.width=7, cache=TRUE}
hospitalflow::admissions_discharges(start_date = start_date, end_date = end_date,
                                    data = example_spell_table,
                                    plot_chart = TRUE,
                                    hospital_name = "Anytown General Hospital")
```


## Length of Stay

```{r los-dist, echo=F, results='asis', fig.height=5, fig.width=7, cache=TRUE}
hospitalflow::los_distrib_method_admission(start_date = start_date, end_date = end_date,
                             data = example_spell_table,
                             plot_chart = TRUE,
                             hospital_name = "Anytown General Hospital")
```
